---
title: "Homework_25"
format: html
---

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
```

```{r}
daphnia<- read.csv(here("Data", "Daphniagrowth.csv"))
```

```{r}
glimpse(daphnia)
daphnia$parasite<- as.factor(daphnia$parasite)
```

```{r}
#way of looking variation in response variable as a function of a grouping categorical variable
ggplot(daphnia, aes(x=parasite, y=growth.rate))+
  geom_boxplot()+
  theme_bw()+
  coord_flip()
#coord_flip makes names easier to read
#shows sig. variaiton in growth rates across treatments 
```

There appears to be a likely overall parasite effect and an ordering of the effect of parasite type on growth rate such that P. ramosa \< M. bicuspidata \< P. perplexa \< control

We can also estimate the average growth rate for each treatment: - control approx 1.2, P. perplexa c. 1.1, M. bicuspidata c. 0.8 and P. ramosa c. 0.55.

-   we can also figure out the degrees of freedom for the treatment and for error. Sample size minus the number of parameters we are estimating = 4 levels - 1 = 3 for the effect of parasites and 4 means from 40 samples = 36 for the error degrees of freedom.


```{r}
#same structure as linear model
model_grow<- lm(growth.rate ~parasite, data = daphnia)
```

Checking assumptions with plot
```{r}
autoplot(model_grow, smooth.colour = NA)
```

upper left
- residuals are fine
upper right
- normality is within the expected bounds of variation
bottom left
- residuals are good
bottom right
- leverages are good

```{r}
anova(model_grow)
#large F and small p allows us to reject the null hypothesis
```

#contrast treatments
```{r}
summary(model_grow)
#shows in alphabetical order
#control
#m. bis
# p. perp
#p. ramosa
```

```{r}
#means for each treatments to contrast ourselves
sumDat<- daphnia |>
  group_by(parasite)|>
  summarise(meanGR = mean(growth.rate))
sumDat
```

```{r}
ggplot(daphnia, aes(parasite, growth.rate, color = parasite))+
     geom_point(size = 3)+
     geom_point(data = sumDat, aes(parasite, meanGR, fill = parasite), shape = 23, size = 5)+
     coord_flip()+
     theme_bw()
```


There is a statistically significant effect of parasite treatment on growth rate of daphnia. (Table 1, F(3,36(d.f)), pvalue<0.05)

f= between/within
MSE parasite= variance between groups
MSE residuals = variance within the groups
