---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("Data", "CrabAgePrediction.csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex<- as.factor(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.
```{r}
crabs_f<- crabs |> filter(Sex == "F")
crabs_f
```


2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:
- Response Weight is numeric and continuous and height the predictor is also numeric and continuous. The data is ration or interval and is continuous and there is a relationship between the two and not more than one predictor variable so a linear regression should be run. 


3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

```{r}
ggplot(crabs_f, aes(x=Height, y=Weight))+
  geom_point()
```

```{r}
model_crabs <- lm(Weight ~ Height, data = crabs_f)
```

```{r}
autoplot(model_crabs, smooth.colour = NA)
```

```{r}
anova(model_crabs)
```

```{r}
summary(model_crabs)
```

```{r}
ggplot(crabs_f, aes(x=Height, y=Weight))+
  geom_point()+
  geom_smooth(method = 'lm')
```

```{r}
ggplot(crabs_f, aes(x=Height, y= Weight ))+
  geom_col()
```

Reject the null hypothesis that there is no association between female crab weight and height. 