---
title: "Homework_26"
format: html
---

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
library(multcomp)
```

```{r}
pea<- read.csv(here("Data", "s_r_pea_sections.csv"))
glimpse(pea)
pea$Treatment<- as.factor(pea$Treatment)
```

```{r}
ggplot(pea, aes(x=Treatment, y=pea_section_length_ou))+
  geom_boxplot()+
  geom_jitter(width = 0.01)+
  theme_bw()+
  coord_flip()
```

- based on the graphs I think there with be a significant difference between control and all the treatments and the between 2 percent sucrose and the other treatments
control mean= 72
2% sucrose = 65
2% glucose= 60
2% fructose = 57
1% fructose 1% glucose = 56

```{r}
 model_pea<- lm(pea_section_length_ou~Treatment, data = pea)
```

```{r}
autoplot(model_pea)
```

- the normality looks okay
- in the leverage and equal variance there seems to be one group that has much larger variance and leverage

```{r}
anova(model_pea)
```

- p-value is significant so there is a difference between at least one of the groups (Table 1, F(4,45)=49.368, p-value< 0.5) so now I will do a tukey HSD to see between which groups there is a difference

```{r}
summary(model_pea)
```

- can see that there is a significant difference between the 1% fructose and sucrose and 2% fructose, 2% sucrose, and control.

```{r}
pea |>
  group_by(Treatment)|>
  summarise(meanGR = mean(pea_section_length_ou))
#mean pea section length for each treatment
```

```{r}
phc1<- glht(model_pea, linfct = mcp(Treatment = "Tukey"))
summary(phc1)
```

```{r}
cld(phc1)
```

- there is a significant difference between 1% fructose and glucose and 2% sucrose and the control, 2% fructose and 2% sucrose and the control, 2% glucose and 2% sucrose and the control, 2% sucrose and control. 


