---
title: "Homework_23"
format: html
---

```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
plant_gr<- read.csv(here("Data", "plant.growth.rate.csv"))
```

Linear Models
Plot-> Model -> Check assumptions -> Interpret -> Plot Again

## Simple Linear Regression
- need a clear relationship between two variables that is easily visualized by plotting the dependent against the independent variable 
- explanatory variable is continuous, numeric variables and does not have categories 

```{r}
glimpse(plant_gr)
```

```{r}
#visualize data
ggplot(plant_gr,
       aes(x = soil.moisture.content, y = plant.growth.rate))+
  geom_point()+
  ylab("Plant Growth Rate (mm/week")+
  theme_bw()
```

## making a Simple Linear Regression Happen
```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content,
                data = plant_gr)
```

```{r}
#diagnostic plots
autoplot(model_pgr, smooth.colour = NA)
#top left whether a line is appropriate to fit the data
#top right assumption of normality of residuals 
#bottom left assumption of equal variance 
#bottom right evaluates leverage to detect influential data points
```

## Now the Interpretation 
```{r}
anova(model_pgr)
#large F-value is large which indicated a small relative variance attributed to the explanatory variable 
summary(model_pgr)
#estimates in a linear regression
```

## ggplot2: the capacity to add regression lines 
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y= plant.growth.rate))+
  geom_point()+
  geom_smooth(method = 'lm')+
  ylab("Plant Growth Rate (mm/week)")+
  theme_bw()
#shove a linear model fitted line and the standard error of the fit using flash transparent

```

