---
title: "13-Nov-2024-NAs"
format: html
editor: visual
---

## Set up our workspace and read in the data

```{r}
rm(list = ls())
library(tidyverse)
library(ggfortify)
library(here)
```

Read in the data.

```{r}
df <- read.csv(here("Data/learn-about-NA.csv"))
str(df)
```


```{r}
df$Treatment<- as.factor(df$Treatment)
levels(df$Treatment)
```

## Dealing with NA's

For your capstone project, you are working with datasets that, most likely, have NA values. How do we deal with them?

### Finding complete cases

In R-speak, a "complete case" is a row of data for which there are NO NA values in ANY of the columns.

Use the `head()` function to show the first 6 rows of `df`. How many complete cases do you see?
```{r}
head(df)
```

**Answer here:**
- 3 complete cases 

The base R function `complete.cases()` is a function that returns boolean values: TRUE for rows that have complete cases and FALSE for rows that have NA in at least 1 column.

Let's use it:

```{r}
complete.cases(df)
```

Now see if you can use `complete.cases()` to check your answer to question 1:

```{r}
complete.cases(head(df))
```

### Counting NA Values

Sometimes it would be nice to determine the number of rows with complete cases or the number of NAs in a dataframe or vector.

Because `complete.cases` returns either TRUE or FALSE, we can determine the number of either TRUE or FALSE rows with the `sum()` function. R will add up the number of TRUE values, thus:

```{r}
sum(complete.cases((df))) #add up the number complete rows in the dataframe
```

We see that there are 36 rows that have no NA values in any columns.

Use a combination of `sum()`, `complete.cases()` and `head()` to return a numeric validation for the number you answered for question 1 above.

**Write your code here:**

```{r}
sum(complete.cases(head(df)))
```

We can also use the base R function `is.na()` to count either the number of NAs or the number of rows with values. Like `complete.cases()`, `is.na()` returns TRUE or FALSE. Take a look.

Let's count the number of rows in the Dried_mass_g column of our dataframe that are NA:

```{r}
sum(is.na(df$Dried_mass_g))
```

What if we wanted to know how many rows in the Dried_mass_g column are NOT NA values, i.e. how many rows have a value? We can use the Not operator `!` like this:

```{r}
sum(!is.na(df$Dried_mass_g))
```

Use the same method to determine the number of rows with NAs in the Growth_rate_mm column and the number of rows without NAs in that column:

```{r}
sum(is.na(df$Growth_rate_mm))

sum(!is.na(df$Growth_rate_mm))
```

### Removing NAs

We have several options for removing cases (i.e. **rows**) with NA values, from both base R and from the tidyverse. Let's look at some base R options first because they are very common and then look at some options from the tidyverse.

Regardless of method, we can either: - remove all rows with NA in **any** column - remove all rows with NA in **a specific** column

Which option we choose can have a big impact on our analysis, as we will see below.

#### Base R Options

We can use `complete.cases()`, `na.omit()` or `is.na()` in our approach:

Use `complete.cases()` to remove all rows that contain NA values in ANY column:

```{r}
df_remove_All <- df[complete.cases(df), ] 
#remember how indexing works: [rows, columns]. Thus, this code says to make a new 
#object called df_remove_All by going to df and grabbing [the rows with complete cases, #]. Leaving just a space after the comma tells R to return all columns of the data #frame.

View(df_remove_All)
```

If, instead, we wanted to return a dataframe in which there were no rows with NA values in the Growth_rate_mm column, we could write some fairly complicated indexing like this:

```{r}
df_remove_NA_growth <- df[complete.cases(df[,3]), ] #go into df and grab the rows in which there are complete cases in the 3rd column and keep all the column of data that match

View(df_remove_NA_growth)
```

See if you can use the same approach to write the code that would return a dataframe in which there are no rows with NA values in the Dried_mass_g column but that returns all columns of the dataframe. Call the object df_remove_NA_Mass

**Write your code here:**


```{r}
df_remove_NA_Mass <- df[complete.cases(df[,2]), ]

View(df_remove_NA_Mass)
```

Another base R for getting rid of rows with NA values is `na.omit()`. Let's take a look.

Use `na.omit()` on df and see what result you get. Save the object as df1.

```{r}
df1 <- na.omit(df)
View(df1)
```

#### Tidyverse methods for removing NAs

Let's get rid of a few of the objects we made to start over with a somewhat clean slate:

```{r}
rm(df1, df_remove_All, df_remove_NA_Mass, df_remove_NA_growth)
```

There is a function in the `tidyr` package called `drop_na()` that we can use to get rid of either rows with ANY NA values or rows with NA values in specific columns.

First, let's get rid of rows with NA values in any column:

```{r}
df1 <- df |> drop_na()

View(df1)
```

Now let's get rid of rows with NA values in the Dried_mass_g column:

```{r}
df_remove_NA_Mass <- df |> drop_na(Dried_mass_g)
```

Now you write some code to use `drop_na()` to create a new data frame called `df_remove_NA_growth` that has no NA values in the Growth_rate_mm column:

```{r}
df_remove_NA_growth <- df |> drop_na(Growth_rate_mm)
```

## Why should we care?

Deciding how to treat NA values in your data can have critical impacts on your data analysis.

Let's look at an example. Let's clean up our workspace so that we have two data frames, our original called df_All_NA, including all of the NA data, and a new data set called df_No_NA that is just the complete cases of df.

```{r}
rm(df_remove_NA_growth, df_remove_NA_Mass)

df_All_NA <- df
df_No_NA <- df1

rm(df, df1)
```

What if we wanted to test the hypothesis that there is no effect of fertilizer treatment on plant growth rate?

Our predictor variable type is: treatment: categorical

Our response variable type is: plant growth rate: continuous 

Therefore our statistical method is: ANOVA

We will do our analysis TWICE, once with all of the data (no NA's removed) and a second time with just the complete cases, to see if there is a difference, and, if so, which approach is better.

### First attempt: All the data

#### Plot and Guess

```{r}
ggplot(df_All_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

What do you think?
there seem to be outliers for each treatment level which could be effecting the mean growth rate. There might be a significant p-value for control and level 2 but not between control and level 1 

#### Create a model:

```{r}
grow.mod<-lm(Growth_rate_mm ~ Treatment, data = df_All_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod)
```

Looks pretty good.

#### Interpret results:

```{r}
anova(grow.mod)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?
three treatments minus one so 2 DF.

Why are there 46 DF for the Residuals?
11 rows were removed 
46+ 11 NA + 3 treatments 

How would you write a results statement?
F(2,46)= 3.6937

```{r}
summary(grow.mod)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?
- it does not as Pvalue= 0.0812 it does not indicate significance. 

Does adding 2 doses of fertilizer change the growth rate signficantly from the Control?
- yes the p-value= 0.0108 so it indicates significance.

We won't worry about re-plotting our data right now.

Now let's do it all again, with the complete cases.

### Second Attempt: Complete Cases

#### Plot and Guess

```{r}
ggplot(df_No_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

What do you think?

#### Create a model:

```{r}
grow.mod.no.NA <-lm(Growth_rate_mm ~ Treatment, data = df_No_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod.no.NA)
```

Fit doesn't look as good. Data appear less normally distributed.

#### Interpret results:

```{r}
anova(grow.mod.no.NA)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

Why are there 33 DF for the Residuals?

How would you write a results statement?

```{r}
summary(grow.mod.no.NA)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

Does adding 2 doses of fertilizer change the growth rate signficantly from the Control?

We won't worry about re-plotting our data right now.
